A general and efficient representation of Ancestral Recombination Graphs

### 2023-08-24 paper.tex
Abstract pass

% FIXME this is weak, but gives the basic outline
\begin{abstract}
Recent breakthroughs have made it possible to infer genetic genealogies in
the presence of recombination at scale, enabling many
downstream applications in population and statistical genetics.
The structure representing such recombinant genetic ancestry
is usually referred to as an Ancestral Recombination Graph (ARG),
although there is some confusion about the interpretation and little
agreement on the concrete details.
We propose a concrete definition of ARGs that is suitable for modern
datasets, in terms of genomes and their patterns of inheritance,
and contrast this with the classical event-based definition which
makes several limiting assumptions.
We show that this new definition enables fine-grained
levels of precision about ancestral processes to be represented,
and forms the basis of an efficient computational framework.
\end{abstract}

>>>>

% FIXME this is weak, but gives the basic outline
\begin{abstract}
Recent breakthroughs have made it possible to infer genetic genealogies in
the presence of recombination at scale, enabling many
downstream applications in population and statistical genetics.
The structure representing such recombinant genetic ancestry
is usually referred to as an Ancestral Recombination Graph (ARG),
although there is some confusion about the interpretation and little
agreement on the concrete details.
We propose a concrete definition of ARGs
in terms of genomes and their intervals of genetic inheritance (gARG)
and contrast this with the classical event-based definition (eARG).
The eARG definition is based on the coalescent with recombination
stochastic process, and we show how the assumptions of that process
place limitations on the patterns of inheritance that can be
represented, and that these assumptions are
% Better wording
substantially broken
in modern large-scale datasets.
Representing genetic ancestry
with an eARG also requires complete precision about the ordering
of all recombination events, which we will rarely have sufficient
information to infer.
We show, in contrast, that the gARG definition
can fully capture the richness of modern large-scale datasets,
enables fine-grained
levels of precision about recombination to be represented,
and forms the basis of an efficient computational framework.
\end{abstract}
